Gravitational Lens Modelling in a Citizen Science Context

### 2014-10-08 text/ms.tex
edit corrections following comments by SM

The encouraging early results from the first \sw lens search, carried
out on the $\simeq172$ square degree Canada-France-Hawaii Telescope
Legacy Survey (CFHTLS) imaging by over 30,000 volunteers (Marshall et
al, in prep; More et al, in prep) prompted the question: could the
modelling of the lenses also be done by the volunteers?  If so,
modelling could help aprioritise lens candidates at an early stage,
which would be very useful with new wide-field and sensitive surveys,
which will yield thousands of lens candidates.  There are several
software tools for lens modelling available, and work has been done on
generic interfaces \citep[e.g.][]{2014A&C.....5...28L}.  Some early
designs for \sw included a prototype lens modelling tool
\citep{2010AAS...21543527N}. Moreover, some \sw volunteers are quite
experienced from earlier projects, having individually spent a
thousand hours or more with data, and are very interested in more
demanding projects.  The interests of citizen-science communities are
just beginning to be studied \citep[e.g.,][]{2013AEdRv..12a0106J}, but
it is clear that some volunteers welcome open-ended challenges, and
sometimes these have led to new scientific results: one example is the
discovery of an exceptional extra-solar planet
\citep{2013ApJ...768..127S}; another is the development of new
algorithms for protein folding \citep{Khatib22112011}.  All these are
grounds for optimism.  There is, however, a basic difficulty in strong
gravitational lensing. Lensed images do not look much like their
source, and still less do they resemble the lensing-mass distribution.
To model a lens, one needs either to do a lot of random guessing, or
to have a good intuition for what works.

>>>>

The encouraging early results from the first \sw lens search, carried
out on the $\simeq172$ square degree Canada-France-Hawaii Telescope
Legacy Survey (CFHTLS) imaging by over 30,000 volunteers (Marshall et
al, in prep; More et al, in prep) prompted the question: could the
modelling of the lenses also be done by the volunteers?  If so,
modelling could help prioritise lens candidates at an early stage,
which would be very useful with new wide-field and sensitive surveys,
which will yield thousands of lens candidates.  There are several
software tools for lens modelling available, and work has been done on
generic interfaces \citep[e.g.][]{2014A&C.....5...28L}.  Some early
designs for \sw included a prototype lens modelling tool
\citep{2010AAS...21543527N}. Moreover, some \sw volunteers are quite
experienced from earlier projects, having individually spent a
thousand hours or more with data, and are very interested in more
demanding projects.  The interests of citizen-science communities are
just beginning to be studied \citep[e.g.,][]{2013AEdRv..12a0106J}, but
it is clear that some volunteers welcome open-ended challenges, and
sometimes these have led to new scientific results: one example is the
discovery of an exceptional extra-solar planet
\citep{2013ApJ...768..127S}; another is the development of new
algorithms for protein folding \citep{Khatib22112011}.  All these are
grounds for optimism.  There is, however, a basic difficulty in strong
gravitational lensing. Lensed images do not look much like their
source, and still less do they resemble the lensing-mass distribution.
To model a lens, one needs either to do a lot of random guessing, or
to have a good intuition for what works.
